Transformation of forecasts for evaluating predictive performance in an epidemiological context

### 2023-01-09 main.tex
update manuscript, address suggestions, delete paragraph

\maketitle
%\tableofcontents
\begin{abstract}
Forecast evaluation plays an essential role in improving forecasting models can help inform future decision making. In epidemiology, forecasts are often scored using the Weighted Interval Score (WIS) or the Continuous Ranked Probability Score (CRPS), which are a measure of absolute distance between forecast and observation. Infectious disease processes are commonly described in terms of the growth rate or the effective reproduction number. Scoring based on the absolute distance between forecast and observation may not reflect the true underlying modelling task well (estimating the growth rate or reproduction number) and makes it difficult to compare forecasts across time, space, and different targets. As a solution we propose transforming the data before scoring the forecasts. Depending on the context, many transformations may be appropriate. We describe the log-transformation as one specific example in detail and discuss its interpretation as a measure of relative error and as score for an implicit forecast of the exponential growth rate. Applying the log-transformation to data from the European COVID-19 Forecast Hub, we found that it changed model rankings, evened scores across time, locations and target types, and emphasised situations in which models missed lower inflection points and consequent rises in numbers, while penalising missed peaks less severely. 
\end{abstract}

@@

Probabilistic forecasts are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data. Proper scoring rules are constructed such that forecasters are incentivised to report their true belief about the future. Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep{bracherEvaluatingEpidemicForecasts2021}. 
The CRPS measures the distance of the predictive distribution to the observed data as 
\begin{linenomath*}
\begin{equation*}
    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,
\end{equation*}    
\end{linenomath*}
where $y$ is the true observed value and $F$ the cumulative distribution function (CDF) of predictive distribution. Both the WIS and the CRPS can be understood as a generalisation of the absolute error to predictive distributions. The WIS is itself an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}. The WIS is commonly defined as 
\begin{linenomath*}
\begin{equation*}
    \text{WIS}(F, y) = \frac{1}{K} \times \sum_{k = 1}^{K} \times 2 \left[ 1(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
\end{equation*}
\end{linenomath*}
where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, and $K$ is the number of predictive quantiles provided.

The dynamics of infectious processes are often described by the complementary concepts of the reproduction number R \citep{gosticPracticalConsiderationsMeasuring2020} and growth rate r \citep{wallingaHowGenerationIntervals2007}, where R describes the strength and r the speed of epidemic growth \citep{dushoffSpeedStrengthEpidemic2021}.
In the absence of changes in immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
In that case, the number of new infections in the population grows exponentially in time.
This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.
If case numbers are growing exponentially and the true modelling task revolves around estimating the reproduction number or the corresponding growth rate, evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount, as errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
If one is to assess the ability of forecasts to assess and forecast the underlying infection dynamics, it may be more appropriate to evaluate errors on the growth rate directly.

Moreover, if forecasts are evaluated based on absolute rather than relative errors, whereas in reality errors scale with the level of observations, scores are strongly influenced by the overall order of magnitude of the quantity to forecast. There exist other scores that do not depend on the scale of the forecast target, such as the log score \citep{goodRationalDecisions1952}, which in addition is also invariant under monotonic transformations of the target, or a scaled version of the CRPS \citep{bolinLocalScaleInvariance2021}. These, however, may not have all of the desired properties of the CRPS and in particular cannot easily be transferred to a setting with predictive quantiles. With the regular CRPS and WIS the dependence on the scale of the forecast target makes it difficult to compare predictive performance across time and in different phases of an epidemic or to compare performance across targets that generally occur at different orders of magnitudes.
In the example of COVID-19 where case numbers underwent drastic changes over time following the initial emergence, forecast evaluation based on the absolute numbers would be dominated by performance around the peaks. The ability to correctly predict future upswings while case numbers are still low would count little towards overall model scores. Scores based on jointly evaluating the ability to predict numbers of cases, hospitalisation and deaths would be dominated by cases as they occur at largest numbers.

In many fields it is common practice to forecast transformed quantities (see e.g. \cite{taylorEvaluatingVolatilityInterval1999} in finance, \cite{mayrLogLevelVAR2015} in macroeconomics, \cite{loweStochasticRainfallrunoffForecasting2014} in hydrology or \cite{fuglstadDoesNonstationarySpatial2015} in meteorology). The goal of the transformations is usually to improve accuracy of the predictions. In this paper, we discuss how transforming the forecasts and observations prior to applying the WIS (or CRPS) may be an appropriate way to improve the evaluation process by addressing the aforementioned issues. Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects of the forecasts that forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}). For conceptual clarity and to allow for a more in-depth discussion, we focus most of our attention on the log-transformation as one example of many possible transformations, chosen for its simplicity and usefulness in an epidemiological context. We further focus on the CRPS/WIS as evaluation metric, given its widespread use throughout the COVID-19 pandemic.

We first outline the motivation for transforming forecasts prior to scoring and specifically the motivation for applying a log-transformation (Section \ref{sec:methods}). We provide a mathematical intuition of the impact of the log-transformation, its interpretation (Section \ref{sec:methods:relative} and \ref{sec:methods:growthrate}) and effects on forecast rankings (section \ref{sec:methods:rankings} and provide some discuss some practical considerations relevant for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}). We illustrate the effect of the log-transformation using forecasts submitted to the European COVID-19 Forecast Hub  \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022} (Section \ref{sec:results}). Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues on the use of transformed scores in the evaluation of infectious disease forecasts (Section \ref{sec:discussion}).

@@

Our aim in transforming forecasts and observations is to shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score. A particularly relevant transformation in the context of epidemic phenomena is the natural logarithm. Instead of a score representing the magnitude of absolute errors this approach makes it possible to obtain a score which a) measures relative error and at the same time b) provides an a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate} and is c) independent from the mean of the observed quantity (see Section \ref{sec:methods:vst}). The interpretation of the log-transformed score as a score for a forecast of the exponential growth rate of a given quantity aligns the evaluation with our understanding of how epidemiological processes should be modeled and the kinds of errors models would typically make.

@@

To illustrate the effect of applying the logarithm prior to evaluating forecasts we consider the absolute error, which the CRPS and WIS generalize to probabilistic forecasts. When considering a point forecast $\hat{y}$ for a quantity of interest $y$, such that 
%
\begin{linenomath*}
\begin{equation*}
y = \hat{y} + \varepsilon,
\end{equation*}
\end{linenomath*}
the absolute error is given by $|\varepsilon|$. When taking the logarithm of the forecast and the observation first (assuming that both are strictly positive), thus considering 
\begin{linenomath*}
\begin{equation*}
\log y = \log \hat{y} + \varepsilon^*,
\end{equation*}
\end{linenomath*}
the resulting absolute error $\left|\varepsilon^*\right|$ can be interpreted as an approximation of various common relative error measures. Using that $\log(a) \approx a - 1$ if $a$ is close to 1, we get
$$
|\varepsilon^*| = |\log \hat{y} - \log y| = \left|\log\left(\frac{\hat{y}}{y}\right) \right| \ \ \stackrel{\text{if } \hat{y} \ \approx \ y}{\approx} \ \ \left| \frac{\hat{y}}{y} - 1 \right| \ \ = \ \ \left| \frac{\hat{y} - y}{y} \right|.
$$
The absolute error after log transformation is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a}. As in the approximation we assumed that $\hat{y} \approx y$, we can just as well interpret it as an approximation of the \textit{relative error} (RE)
$$
\left| \frac{\hat{y} - y}{\hat{y}} \right|
$$
and the \textit{symmetric absolute percentage error} (SAPE)
$$
\left| \ \frac{\hat{y} - y}{y/2 + \hat{y}/2} \ \right|.
$$
As Figure \ref{fig:SAPE} shows, the alignment with the SAPE is in fact the closest and holds quite well even if predicted and observed value differ by a factor of two or three. Moving back to probabilistic forecasts, the CRPS or WIS applied to log-transformed forecasts and outcomes can thus be seen as a probabilistic counterpart to the symmetric absolute percentage error, which offers an appealing intuitive interpretation.

\begin{figure}[h!]
\centering
\includegraphics[width = 1\textwidth]{output/figures/different-relative-errors.png}
\caption{Numerical comparison of different measures of relative error/ We consider a prediction of $\hat{y} = 1$ and for different observed values show the  absolute percentage error (APE), relative error (RE), symmetric absolute percentage error (SAPE) and the absolute error applied to log-transformed predictions and observations. $\hat{y}$ denotes the predicted value. A: Error measures as a function of the observed value with the x-axis shown on a linear scale. B: Same, but x-axis shown on a logarithmic scale.}
\label{fig:SAPE}
\end{figure}

@@

Another interpretation can be based around viewing the process to be forecast as exponential with time-varying growth rate $r(t)$~\citep[see, e.g.,][]{wallingaHowGenerationIntervals2007}, i.e.
\begin{linenomath*}
\begin{equation*}
\frac{d}{dt}y(t) = r(t)y(t)
\end{equation*}
\end{linenomath*}
%
which is solved by
%
\begin{linenomath*}
\begin{equation*}
y(t) = y_0 \exp \left( \int_0^t r(t') dt' \right) = y_0 \exp (\bar{r}t)
\end{equation*}
\end{linenomath*}
where $y_0$ is an initial data point and $\bar{r}$ is the mean of the growth rate between the initial time point $0$ and time $t$.

If a forecast $\hat{y}(t)$ for the value of the time series at time $t$ is issued at time 0 based on the data point $y_0$ then the absolute error after log transformation is
%
\begin{linenomath*}
\begin{align*}
\epsilon^* &= \left| \log \left[ \hat{y}( t ) \right] - \log \left[ y ( t ) \right] \right|\\
           &= \left| \log \left[ y_0 \exp (\bar{\hat{r}} t ) \right] - \log \left[ y_0 \exp (\bar{r}t) \right] \right|\\
           &= t \left| \bar{\hat{r}} - \bar{r} \right|.
\end{align*}
\end{linenomath*}
%
We thus evaluate the error in the mean exponential growth rate, scaled by the length of the time period considered. For small $\bar{r}t$, $ y_0 \exp (\bar{r}t)$ is approximately equivalent to $ y_0 (1 + \bar{r}t)$, which is sometimes used as an alternative definition of the growth rate (then called the multiplicative growth rate).

@@

In epidemiological contexts, the variance of the variable of interest usually scales with its mean.  A variance-stabilising transformation \citep[VST][]{bartlettSquareRootTransformation1936} is a transformation that serves to make the variance of a variable independent of its mean.

If the predictive distribution for the quantity $Y$ equals the true data-generating process $F$ (an ideal forecast), the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
$$
\mathbb{E}[\text{CRPS}(F, y)] = 0.5\times\mathbb{E}|Y - Y'|,
$$
where $Y$ and $Y'$ are independent samples from $F$. This corresponds to half the \textit{mean absolute difference}, which is a measure of dispersion. If $F$ is well-approximated by a normal distribution $\text{N}(\mu, \sigma^2)$, the approximation
$$
\mathbb{E}_F[\text{CRPS}(F, y)] \approx \frac{\sigma}{\sqrt{\pi}}
$$
can be used, meaning that the expected CRPS scales roughly with the standard deviation. Denoting by $F^*$ the predictive distribution for $\log(Y)$, the delta method implies that
$$
\mathbb{E}_F[\text{CRPS}\{F^*, \log(y)\}] \approx \frac{\sigma/\mu}{\sqrt{\pi}}.
$$
The expected CRPS for the log-transformed outcome thus scales with the coefficient of variation. In instances where the variance scales with the square of the mean, the log-transformation serves as a variance-stabilising transformation \citep{guerreroTimeseriesAnalysisSupported1993}, making the expected CRPS independent of the mean of the forecast quantity. We illustrate this in Figure \ref{fig:SIM-wis-state-size-mean}, where we show CRPS scores for different ideal forecasters. For a negative-binomial distribution with $\sigma^2 = \mu + \mu^2$ and sufficiently large $\mu$ we see that on the log scale scores are independent of the mean of the forecast target. For a Poisson-distribution with 
$\sigma = \mu$, scores on the log scale decrease as the mean increases. The appropriate variance-stabilising transformation here to make scores independent of the mean would be a square-root-transformation \citep{bartlettSquareRootTransformation1936}.

\subsection{Effects on model rankings}
\label{sec:methods:rankings}
Rankings between different forecasters based on their scores may change through the log-transformation both when looking at aggregate scores as well as when looking at a single forecast. We illustrate this in Figure \ref{fig:illustration-ranking} with two forecasters, A and B, issuing two different distributions with different dispersion. When showing the obtained score as a function of the observed value, it can be seen that the ranking between the two forecasters may change when scoring the forecast on the logarithmic, rather than the natural scale. In particular, on the natural scale, forecaster A, who issues a more uncertain distribution, receives a better score than forecaster B for observed values far away from the centre of the respective predictive distribution. On the log scale, however, forecaster A receives a lower score for large observed values, being more heavily penalised for assigning large probability to small values (which, in relative terms, are far away from the actual observation).

@@

Overall model rankings would be expected to differ even more when scores are averaged across multiple forecasts or targets. The change in rankings of aggregate scores is not mainly driven by the kinds of changes in model rankings for single forecasts just discussed. Rather, as observations from exponential processes are likely to be on very different absolute scales, large observations will account for the majority of absolute score but will not dominate relative scores in the same way. Evaluating forecasts based on relative errors may help with comparing performance across different targets. However, log-transforming forecasts and observations may also result in the opposite extreme because small absolute errors (e.g. predicting 9, rather than 3 deaths) may be large in relative terms if the quantity to forecast is very small. This can be exacerbated when forecasts are restricted to discrete values, where relative steps between possible values become greater for values closer to zero.

@@

\begin{equation*}
    \text{CRPS}(\log F, \log y) = \alpha^* + \beta^* \cdot \log{(\text{median}(F))}. 
\end{equation*}

>>>>

\maketitle
%\tableofcontents
\begin{abstract}
Forecast evaluation plays an essential role in improving forecasting models can help inform future decision making. In epidemiology, probabilistic forecasts are often scored using the Weighted Interval Score (WIS) or the Continuous Ranked Probability Score (CRPS), which are a measure of absolute distance between forecast and observation. Infectious disease transmission is commonly conceptualised as a non-stationary processes in terms of the growth rate or the effective reproduction number. Scoring based on the absolute distance between forecast and observation may not reflect this underlying modelling task well and may make it difficult to compare forecasts across time, space, and different forecast targets. As a solution, we propose transforming the data before scoring the forecasts to better reflect our understanding of the generative process. Depending on the context, many transformations may be appropriate. We describe the log-transformation as one specific example in detail and discuss its interpretation as a measure of relative error and as a score for an implicit forecast of the exponential growth rate. Applying the log-transformation to data from the European COVID-19 Forecast Hub, we found that it changed model rankings, regardless of stratification by forecast date, location or target types, and emphasised situations in which models missed lower inflection points and consequent rises in numbers, while penalising missed peaks less severely. 
\end{abstract}

@@

Probabilistic forecasts are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data. Proper scoring rules are constructed such that forecasters are incentivised to report their true belief about the future. Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep{bracherEvaluatingEpidemicForecasts2021}. 
The CRPS measures the distance of the predictive distribution to the observed data as 
\begin{linenomath*}
\begin{equation*}
    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,
\end{equation*}    
\end{linenomath*}
where $y$ is the true observed value and $F$ the cumulative distribution function (CDF) of the predictive distribution. The CRPS can be understood as a generalisation of the absolute error to predictive distributions. The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}. The WIS is defined as 
\begin{linenomath*}
\begin{equation*}
    \text{WIS}(F, y) = \frac{1}{K} \times \sum_{k = 1}^{K} 2 \times \left[ \boldsymbol{1}(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
\end{equation*}
\end{linenomath*}
where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of predictive quantiles provided and $\boldsymbol{1}$ is the indicator function. We show an alternative definition based on central prediction intervals in Section \ref{sec:alternative-wis}.

The dynamics of infectious processes are often described by the complementary concepts of the reproduction number $R$ \citep{gosticPracticalConsiderationsMeasuring2020} and growth rate $r$ \citep{wallingaHowGenerationIntervals2007}, where $R$ describes the strength and $r$ the speed of epidemic growth \citep{dushoffSpeedStrengthEpidemic2021}.
In the absence of changes in immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
In that case, the number of new infections in the population grows exponentially in time.
This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.
If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating the reproduction number or the corresponding growth rate, evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount, as errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
If one is to measure the ability of forecasts to assess and forecast the underlying infection dynamics, it may thus be more desirable to evaluate errors on the growth rate directly.

Evaluating forecasts using the CRPS or WIS means that scores represent a measure of absolute errors. However, forecast consumers may find errors on a relative scale easier to interpret and more useful in order to track predictive performance across targets of different orders of magnitude. 
\cite{bolinLocalScaleInvariance2021} have proposed the scaled CRPS (SCRPS) which is locally scale invariant; however, it does not correspond to a relative error measure and lacks a straightforward interpretation as available for the CRPS. % is less commonly used and does not have a straightforward equivalent for a setting with predictive quantiles. 
A closely related aspect to relative scores (as opposed to absolute scores) is that in the evaluation one may wish to give similar weight to all considered forecast targets. As the CRPS typically scales with the order of magnitude of the quantity to be predicted, this is not the case for the CRPS, which will typically assign higher scores to forecast targets with high expected values (e.g., in large locations or around the peak of an epidemic). \cite{bracherEvaluatingEpidemicForecasts2021} have argued that this is a desirable feature, directing attention to situations of particular public health relevance. An evaluation based on absolute errors, however, will assign little weight to other potentially important aspects, such as the ability to correctly predict future upswings while observed numbers are still low.

%that do not depend on the scale of the forecast target, such as the log score \citep{goodRationalDecisions1952}, which in addition is also invariant under monotonic transformations of the target, or a scaled version of the CRPS \citep{bolinLocalScaleInvariance2021}. These, however, may not have all of the desired properties of the CRPS and in particular cannot easily be transferred to a setting with predictive quantiles.

%In this paper we argue that these three aspects (evaluating the growth rate rather than absolute values, evaluating relative errors, giving more equal weights to all forecast targets) can all be addressed by simply applying the natural logarithm to forecasts (in our case predictive quantiles) and observations prior to computing the CRPS.

In many fields it is common practice to forecast transformed quantities (see e.g. \cite{taylorEvaluatingVolatilityInterval1999} in finance, \cite{mayrLogLevelVAR2015} in macroeconomics, \cite{loweStochasticRainfallrunoffForecasting2014} in hydrology or \cite{fuglstadDoesNonstationarySpatial2015} in meteorology). While the goal of the transformations is usually to improve accuracy of the predictions, they can also be used to enhance and complement the evaluation process. 
In this paper we argue that the aforementioned issues with evaluating forecasts based on measures of absolute error can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation. Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score. Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects of the forecasts that forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).

For conceptual clarity and to allow for a more in-depth discussion, we focus mostly on the natural logarithm as a particular transformation in the context of epidemic phenomena. Instead of a score representing the magnitude of absolute errors, applying a log-transformation prior to the CRPS yields a score which a) measures relative error (see Section \ref{sec:methods:relative}), b) provides a measure for how well a forecast captures the exponential growth rate of the target quantity (see Section \ref{sec:methods:growthrate}) and c) is less dependent on the expected order of magnitude of the quantity to be predicted (see Section \ref{sec:methods:vst}). %The interpretation of the log-transformed score as a score for a forecast of the exponential growth rate of a given quantity aligns the evaluation with our understanding of how epidemiological processes should be modeled and the kinds of errors models would typically make. 
We therefore argue that such evaluations on the logarithmic scale should complement the prevailing evaluations on the natural scale. 
Other transformations may likewise be of interest, and we briefly explore the square root transformation as an alternative transformation. %[and population standardization?]. 
Our analysis focuses on the CRPS (or WIS) as evaluation metric, given its widespread use throughout the COVID-19 pandemic.

The remainder of the article is structured as follows. In Sections \ref{sec:methods:relative}--\ref{sec:methods:vst} we provide some mathematical intuition on applying the log transformation prior to evaluating the CRPS, highlighting the connections to relative error measures, the epidmemic growth rate and variance stabilizing transformations.
%We first provide a mathematical intuition of the impact of the log-transformation and its interpretation as relative error (Section \ref{sec:methods:relative}, measure of the accuracy of a forecast of the growth rate (Section \ref{sec:methods:growthrate}) and it's effect as a variance-stabilising transformation that makes can make scores independent of the mean of the forecast target (Section \ref{sec:methods:vst}. 
We then discuss the effect of the transformation on forecast rankings (section \ref{sec:methods:rankings} and discuss some practical considerations relevant for applying transformations in general and the log-transformation in particular (Section \ref{sec:methods:considerations}). We illustrate the effect of the log-transformation using forecasts submitted to the European COVID-19 Forecast Hub  \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022} (Section \ref{sec:results}). Finally, we provide scoring recommendations, discuss alternative transformations that may be useful in different contexts, and suggest further research avenues on the use of transformed scores in the evaluation of infectious disease forecasts (Section \ref{sec:discussion}).

@@

@@

To illustrate the effect of applying the logarithm prior to evaluating forecasts we consider the absolute error, which the CRPS and WIS generalize to probabilistic forecasts. When considering a point forecast $\hat{y}$ for a quantity of interest $y$, such that 
%
\begin{linenomath*}
\begin{equation*}
y = \hat{y} + \varepsilon,
\end{equation*}
\end{linenomath*}
the absolute error is given by $|\varepsilon|$. When taking the logarithm of the forecast and the observation first (assuming that both are strictly positive), thus considering 
\begin{linenomath*}
\begin{equation*}
\log y = \log \hat{y} + \varepsilon^*,
\end{equation*}
\end{linenomath*}
the resulting absolute error $\left|\varepsilon^*\right|$ can be interpreted as an approximation of various common relative error measures. Using that $\log(a) \approx a - 1$ if $a$ is close to 1, we get
$$
|\varepsilon^*| = |\log \hat{y} - \log y| = \left|\log\left(\frac{\hat{y}}{y}\right) \right| \ \ \stackrel{\text{if } \hat{y} \ \approx \ y}{\approx} \ \ \left| \frac{\hat{y}}{y} - 1 \right| \ \ = \ \ \left| \frac{\hat{y} - y}{y} \right|.
$$
The absolute error after log transformation is thus an approximation of the \textit{absolute percentage error} \citep[APE,][]{gneitingMakingEvaluatingPoint2011a}. As in the approximation we assumed that $\hat{y} \approx y$, we can just as well interpret it as an approximation of the \textit{relative error} (RE)
$$
\left| \frac{\hat{y} - y}{\hat{y}} \right|
$$
and the \textit{symmetric absolute percentage error} (SAPE)
$$
\left| \ \frac{\hat{y} - y}{y/2 + \hat{y}/2} \ \right|.
$$
As Figure \ref{fig:SAPE} shows, the alignment with the SAPE is in fact the closest and holds quite well even if predicted and observed value differ by a factor of two or three. Moving back to probabilistic forecasts, the CRPS applied to log-transformed forecasts and outcomes can thus be seen as a probabilistic counterpart to the symmetric absolute percentage error, which offers an appealing intuitive interpretation.

\begin{figure}[h!]
\centering
\includegraphics[width = 1\textwidth]{output/figures/different-relative-errors.png}
\caption{Numerical comparison of different measures of relative error: absolute percentage error (APE), relative error (RE), symmetric absolute percentage error (SAPE) and the absolute error applied to log-transformed predictions and observations. We denote the predicted value by $\hat{y}$ and display errors as a function of the ratio of observed and predicted value. A: x-axis shown on a linear scale. B: x-axis shown on a logarithmic scale.}
\label{fig:SAPE}
\end{figure}

@@

Another interpretation can be based on viewing the process to be forecast as exponential with time-varying growth rate $r(t)$~\citep[see, e.g.,][]{wallingaHowGenerationIntervals2007}, i.e.
\begin{linenomath*}
\begin{equation*}
\frac{d}{dt}y(t) = r(t)y(t)
\end{equation*}
\end{linenomath*}
%
which is solved by
%
\begin{linenomath*}
\begin{equation*}
y(t) = y_0 \exp \left( \int_0^t r(t') dt' \right) = y_0 \exp (\bar{r}t)
\end{equation*}
\end{linenomath*}
where $y_0$ is an initial data point and $\bar{r}$ is the mean of the growth rate between the initial time point $0$ and time $t$.

If a forecast $\hat{y}(t)$ for the value of the time series at time $t$ is issued at time 0 based on the data point $y_0$ then the absolute error after log transformation is
%
\begin{linenomath*}
\begin{align*}
\epsilon^* &= \left| \log \left[ \hat{y}( t ) \right] - \log \left[ y ( t ) \right] \right|\\
           &= \left| \log \left[ y_0 \exp (\bar{\hat{r}} t ) \right] - \log \left[ y_0 \exp (\bar{r}t) \right] \right|\\
           &= t \left| \bar{\hat{r}} - \bar{r} \right|.
\end{align*}
\end{linenomath*}
%
We thus evaluate the error in the mean exponential growth rate, scaled by the length of the time period considered. For small $\bar{r}t$, $ y_0 \exp (\bar{r}t)$ is approximately equivalent to $ y_0 (1 + \bar{r}t)$, which is sometimes used as an alternative definition of the growth rate (then called the multiplicative growth rate). This means that by applying the CRPS rather than the absolute error we perform a probabilistic evaluation of forecasts of the epidemic growth rate.

@@

When evaluating models across sets of forecasting tasks, it may be considered desirable for each target to have a similar impact on the overall results. In disease forecasting this is not the case when using the CRPS, as the latter typically scales with the order of magnitude of the quantity to be predicted. Average scores are then dominated by the results achieved for targets with high expected outcomes. 
%Bracher et al. mention this as a desirable feature of the CRPS. 
% In epidemiological contexts, the variance of the variable of interest usually scales with its mean. 
Specifically, if the predictive distribution for the quantity $Y$ equals the true data-generating process $F$ (an ideal forecast), the expected CRPS is given by \citep{gneitingStrictlyProperScoring2007}
$$
\mathbb{E}[\text{CRPS}(F, y)] = 0.5\times\mathbb{E}|Y - Y'|,
$$
where $Y$ and $Y'$ are independent samples from $F$. This corresponds to half the \textit{mean absolute difference}, which is a measure of dispersion. If $F$ is well-approximated by a normal distribution $\text{N}(\mu, \sigma^2)$, the approximation
$$
\mathbb{E}_F[\text{CRPS}(F, y)] \approx \frac{\sigma}{\sqrt{\pi}}
$$
can be used. This means that the expected CRPS scales roughly with the standard deviation, which in turn typically increases with the mean in epidemiological forecasting. In order to make the expected CRPS independent of the expected outcome, a \textit{variance-stabilising transformation} \citep[VST,][]{bartlettSquareRootTransformation1936} can be employed. The choice of this transformation depends on the mean-variance relationship of the underlying process.

If the mean-variance relationship is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST \citep{guerreroTimeseriesAnalysisSupported1993}. Denoting by $F^*$ the predictive distribution for $\log(Y)$, we can use the delta method to show that
$$
\mathbb{E}_F[\text{CRPS}\{F^*, \log(y)\}] \approx \frac{\sigma/\mu}{\sqrt{\pi}} 
= \frac{\sqrt{c}}{\sqrt{\pi}}
.
$$
The assumption of a quadratic mean-variance relationship is closely linked to the aspects discussed in Sections \ref{sec:methods:relative} and \ref{sec:methods:growthrate}. It implies that relative errors have constant variance and can thus be meaningfully compared across different targets. Also, it arises naturally if we assume that our capacity to predict the epidemic growth rate does not depend on the expected outcome.

If the variance is linear with $\sigma^2 = c \times \mu$, as with a Poisson-distributed variable, the square root is known to be a VST. Denoting by $\tilde{F}$ the predictive distribution for $\sqrt{Y}$, the delta method can again be used to show that
$$
\mathbb{E}_F[\text{CRPS}\{\tilde{F}, \sqrt{y}\}] \approx \frac{\sigma/\sqrt{\mu}}{2\sqrt{\pi}} = \frac{\sqrt{c}}{2\sqrt{\pi}}
.
$$

% which one is an empirical question - maybe hint that quadratic may result from RW on growth rate?

To strengthen our intuition on how transforming outcomes prior to applying the CRPS shifts the emphasis between targets with high and low expected outcomes, Figure \ref{fig:SIM-wis-state-size-mean} shows the expected CRPS of ideal forecasters under different mean-variance relationships and transformations. We consider a Poisson distribution where $\sigma^2 = \mu$, a geometric distribution where $\sigma^2 = \mu + \mu^2$, and a negative binomial distribution where the size parameter $\psi(\mu)$ depends on the mean $\mu$ such that $\sigma^2 = \mu + \mu^{2.5}/1000$. We see that when applying the CRPS on the natural scale, the expected CRPS grows in $\mu$ in all cases. When applying it after a log-transformation, the expected CRPS is almost independent of $\mu$ for the geometric distribution and large $\mu$, while smaller targets have higher expected CRPS in the Poisson case. When applying a square-root-transformation before the CRPS, the expected CRPS is independent of the mean for the Poisson-distribution and grows for the two other distributions. 
%For the negative binomial distribution, targets with very small and very high expectation yield the highest expected CRPS. \textbf{Here it would be nice to also add a sentence on what happens under sqrt transformation.} % scores are independent of the mean of the forecast target. For a Poisson-distribution with 
% $\sigma = \mu$, scores on the log scale decrease as the mean increases. The appropriate variance-stabilising transformation here to make scores independent of the mean would be a square-root-transformation .

\subsection{Effects on model rankings}
\label{sec:methods:rankings}
Rankings between different forecasters based on the CRPS may change when making use of a transformation both when looking at aggregate scores as well as when looking at a single forecast. We illustrate this in Figure \ref{fig:illustration-ranking} with two forecasters, A and B, issuing two different distributions with different dispersion. When showing the obtained CRPS as a function of the observed value, it can be seen that the ranking between the two forecasters may change when scoring the forecast on the logarithmic, rather than the natural scale. In particular, on the natural scale, forecaster A, who issues a more uncertain distribution, receives a better score than forecaster B for observed values far away from the centre of the respective predictive distribution. On the log scale, however, forecaster A receives a lower score for large observed values, being more heavily penalised for assigning large probability to small values (which, in relative terms, are far away from the actual observation).

@@

Overall model rankings would be expected to differ even more when scores are averaged across multiple forecasts or targets. The change in rankings of aggregate scores is not mainly driven by the kinds of changes in model rankings for single forecasts just discussed. Rather, average scores will be most influenced by the order of magnitude of scores for different forecast targets across time, location and target type. On the natural scale, large observations will account for the majority of the overall score but will not dominate average CRPS values obtained after log transformation in the same way. Depending on the relationship between mean and variance of the forecast target, a log-transformation, however, may also lead to systematically larger scores assigned to small forecast targets, as illustrated in Figure \ref{fig:SIM-wis-state-size-mean}. A different transformation like the square-root may then be appropriate to stabilise the variance effectively and make scores independent of the mean of the target quantity.

@@

\begin{equation*}
    \text{CRPS}(\log F, \log y) = \alpha^* + \beta^* \cdot \log{(\text{median}(F))}
\end{equation*} 
and the following regression on scores obtained after applying a square-root-transformation

\begin{equation*}
    \text{CRPS}(\sqrt{F}, \sqrt{y}) = \alpha^{**} + \beta^{**} \cdot \sqrt{(\text{median}(F))}
\end{equation*}
